Nonlinear system identification using machine learning tools

### 2021-03-14 sysID.tex
write out the model details.

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{amssymb}
%\usepackage{lucidbry}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{subcaption}
\usepackage{physics}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,tikzsettings}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage[english]{babel}
\bibliographystyle{plainnat}
%\setcitestyle{numeric,open={(},close={)}} \usepackage{color}
%\usepackage{colortbl}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{shapes,calc,backgrounds,arrows,fit,decorations.pathmorphing,matrix}
\usepackage[margin=0.79in]{geometry}

@@

We next study the performance of the modeling frameworks on a simple two reaction system example in three steps. First, we examine the predictive capability of the developed models on open-loop validation data. Second, we analyze the steady-state optimum with an economic stage cost. Finally, we analyze the optimal control input profiles on a dynamic economic optimization problem. These three steps shed light on the predictive abilities of the models and their usefulness for real-time optimization in the operation of chemical processes.

@@

Before analyzing a full closed-loop simulation with the economic MPC controllers based on the different dynamic models, we first examine the open-loop solutions of the optimization problem \eqref{eq:dyn_opt} given one initial state. Figure \ref{fig:tworeac_openloop} shows the open-loop measurement and control input profiles for the plant, grey-box, black-box NN, and Koopman operator model. We notice that the black-box NN has an optimal control input profile qualitatively similar to that of the plant model.
Whereas, the optimal profiles for the grey-box and Koopman operator model
are different than the plant model. This simple analysis shows that the use of the grey-box and Koopman operator models in MPC can result in suboptimal performance.

\section{Large, CSTR and Flash system example}

@@

@@

\section{Future work}

>>>>

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{amssymb}
%\usepackage{lucidbry}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{subcaption}
\usepackage{physics}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,tikzsettings}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{float}
\usepackage{units}
\usepackage[round]{natbib}
\usepackage[english]{babel}
\bibliographystyle{plainnat}
%\setcitestyle{numeric,open={(},close={)}} \usepackage{color}
%\usepackage{colortbl}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{shapes,calc,backgrounds,arrows,fit,decorations.pathmorphing,matrix}
\usepackage[margin=0.79in]{geometry}

@@

We now study the performance of the modeling frameworks on a simple two
reaction system example in three steps. First, we examine the predictive
capability of the developed models on open-loop validation data. Second, we
analyze the steady-state optimum with an economic stage cost. Finally, we
analyze the optimal control input profiles on a dynamic economic optimization
problem. These three steps shed light on the predictive abilities of the models
and their usefulness for real-time optimization in the operation of chemical
processes.

@@

Before analyzing a full closed-loop simulation with the economic MPC controllers
based on the different dynamic models, we first examine the open-loop solutions
of the optimization problem \eqref{eq:dyn_opt} given one initial state. Figure
\ref{fig:tworeac_openloop} shows the open-loop state and control input profiles
for the plant, grey-box, black-box NN, and Koopman operator models. We notice
that the black-box NN has an optimal control input profile qualitatively similar
to that of the plant. Whereas, the optimal profiles for the grey-box and Koopman
operator model are different than the plant model. This simple analysis shows
that the use of the grey-box and Koopman operator models in MPC can result in
suboptimal economic performance.

\section{Large, CSTR and flash system example}
We next study the performance of the black-box NN modeling framework on a large chemical process that consists of a chemical reactor and flash separator in series. The process is depicted in Figure \ref{fig:cstr_flash}, and the ordinary differential equations (ODEs) with parameter values describing the plant model are shown in Appendix \ref{app:cstr_flash}.

\begin{figure}[!h]
  \centering
  \includegraphics[page=1, height=0.3\textheight,
                   width=0.6\textwidth]{cstr_flash.pdf}
  \caption{Schematic of the CSTR and flash chemical process.}
  \label{fig:cstr_flash}
\end{figure}

@@

\subsection{Steady state optimization}

@@

\section{Discussion}

\section*{Appendix} 
\renewcommand{\thesubsection}{\Alph{subsection}}

\subsection{CSTR and flash system model} \label{app:cstr_flash}

\begin{align*}
  \dfrac{dH_r}{dt} &= \dfrac{F + D -F_r}{A_r}\\
  \dfrac{dC_{Ar}}{dt} &= \dfrac{F(C_{Af} -C_{Ar}) +
                         D(C_{Ad} -C_{Ar})}{A_rH_r} - r_1 \\
  \dfrac{dC_{Br}}{dt} &= \dfrac{-FC_{Br} + 
                          D(C_{Bd} -C_{Br})}{A_rH_r} + r_1 -3r_2\\
  \dfrac{dC_{Cr}}{dt} &= \dfrac{-FC_{Cr} + 
  D(C_{Cd} -C_{Cr})}{A_rH_r} + r_2\\
  \dfrac{dT_r}{dt} &= \dfrac{F(T_f - T_r) + D(T_d -T_r)}{A_rH_r} + 
                      \dfrac{r_1\Delta H_1 + r_2\Delta H_2}{\rho C_p} - 
                      \dfrac{Q_r}{\rho A_r C_p H_r}\\
\end{align*}

\begin{align*}
  \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
  \dfrac{dC_{Ab}}{dt} &= \dfrac{F_r(C_{Ar} -C_{Ab}) + 
                          D(C_{Ab} -C_{Ad})}{A_bH_b} \\
  \dfrac{dC_{Bb}}{dt} &= \dfrac{F_r(C_{Br} -C_{Bb}) + 
                          D(C_{Bb} -C_{Bd})}{A_bH_b} \\
  \dfrac{dC_{Cb}}{dt} &= \dfrac{F_r(C_{Cr} -C_{Cb}) + 
                          D(C_{Cb} -C_{Cd})}{A_bH_b} \\
  \dfrac{dT_b}{dt} &= \dfrac{F_r(T_r - T_b)}{A_bH_b} +
                      \dfrac{Q_b}{\rho A_b C_p H_b}\\
\end{align*}

Here, the rate of reactions $r_1$ and $r_2$ are defined as:
\begin{align*}
  r_1 = k_1^*e^{-E_1/RT_r}C_{Ar}, \quad r_2 = k_2^*e^{-E_2/RT_r}C_{Br}^3
\end{align*}

The outlet flowrates from the CSTR and flash are computed from the height of the reaction mixture in each unit as follows:
\begin{align*}
  F_r = k_r\sqrt{H_r}, \quad F_b = k_b\sqrt{H_b}
\end{align*}

The concentrations of the species in the recycle stream are computed using the following equations:
\begin{align*}
  C_{Ad} = \dfrac{\alpha_A C_{Ab}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}} , \quad, C_{Bd} = \dfrac{\alpha_B C_{Bb}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}}, \quad C_{Cd} = \dfrac{\alpha_C C_{Cb}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}}
\end{align*}

The parameters in this model and the control input constraints used for the optimization problems are summarized in Table \ref{table:cstr_flash_pars}.

\begin{table}[!h]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|}
      \hline
      \multicolumn{6}{ |c| }{Parameters used to simulate the ODEs} \\
      \hline
      Parameter & Value & Unit & Parameter & Value & Unit\\
      \hline
      $\alpha_A$ & 8 &  & $k_1^*$ & 0.3 & \unit{min$^{-1}$}\\ 
      $\alpha_B$ & 1 &  & $k_2^*$ & 0.5 & \unit{min$^{-1}$} \\ 
      $\alpha_C$ & 1 &  & $\Delta H_1$ & 100 & \unitfrac{kJ}{mol} \\ 
      $\rho$ & 6 & \unitfrac{kg}{m$^3$} & $\Delta H_2$ & 120 &
      \unitfrac{kJ}{mol} \\ 
      $C_p$ & 3 & \unitfrac{kJ}{kg-K} & $E_1/R$ & 200 & \unit{K} \\ 
      $A_r$ & 3 & \unit{m$^2$} & $E_2/R$ & 300 & \unit{K} \\ 
      $A_b$ & 3 & \unit{m$^2$} & $T_d$ & 310 &
      \unit{K} \\ 
      $k_r$ & 4 & \unitfrac{m$^{5/2}$}{min} & $C_{Af}$ & 6 & \unitfrac{mol}{m$^3$} \\ 
      $k_b$ & 2 & \unitfrac{m$^{5/2}$}{min} &  $Q_r$ &  200 & \unitfrac{kJ}{min} \\ 
      $T_f$ & 320 & \unit{K} &  & &  \\ 
      $Q_b$ & 200 & \unitfrac{kJ}{min} & & & \\ 
      \hline       
      \multicolumn{6}{ |c| }{Actuator constraints ($\underline{u},
      \overline{u}$)} \\
      \hline
      $F$ & (5, 15) & \unitfrac{m$^3$}{min} & $D$ & (2, 8) &
      \unitfrac{m$^3$}{min}  \\ 
      \hline
  \end{tabular} \caption{Parameters used in the ODEs to simulate the plant in
  the CSTR in series with a flash separator example, and the actuator constraints
  for the optimization problems.}
\label{table:cstr_flash_pars}
\end{table}
